Viscous Erosion in Porous Media

### 2018-01-09 docs/methods/erosion_methods.tex
worked on intro

\title{Methods paper for erosion}

@@

\begin{keyword}
  Stokes flow \sep Erosion \sep Boundary integral method \sep
  Fluid-structure interaction \sep Fast multipole methods 
\end{keyword}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction\label{s:intro}}
Erosion by flowing fluid plays an essential role in diverse processes.
Many of the flows are geophysical, such as the formation of sand dunes
or erosion due to groundwater flow.  In both these examples, there is a
feedback process between the flow and the erodible body.  In particular,
the shape of the eroding body affects the flow, while the flow changes
the shape of the eroding body.  This fluid-mechanical feedback
system has be simulated in idealized cases such as a single eroding body
in a uniform flow~\cite{moo-ris-chi-zha-she2013}.  To further develop
insight into the effect of two-dimensional erosion, we consider erosion
of a multiply-connected domain (Figure~\ref{fig:50bodies}), such as a
porous media that is typically used to model groundwater flow.

Because of the scales involved with groundwater flow, we will restrict
our attention to Stokesian dynamics.  That is, the fluid equations will
be governed by the incompressible Stokes equations.  Because of the
linearity of the underlying equations, a boundary integral equation is
well-suited as it can resolve complex geometries, the interactions
between bodies, all the while achieving high-order accuracy and
computational efficiency.

\paragraph{Contributions} The present study will develop numerical
methods to simulate viscous erosion of a multiply-connected
two-dimensional geometry.  By assuming the flow is viscous, we are
assuming the flow is very slow and in small length scales, as is the
case in groundwater flow.  Moreover, by allowing for complex
multiply-connected domains, the erosion caused by the non-local complex
interactions between different bodies is modeled.

@@

\paragraph{Limitations} There are two main limitations of the current
work.  First, the flow is assumed to be two-dimensional.  Although
several flows can be described to good accuracy under a two-dimensional
approximation, certain phenomena occur only in three-dimensional.  The
boundary integral equation formulation for the fluid solver exists, and
the main challenges of a BIE solver have been addressed by other groups
including surface representation~\cite{yin-bir-zor2006,
vee-rah-bir-zor2011}, near-singular integration~\cite{yin-bir-zor2006,
kli-tor2016b}, and fast summation methods~\cite{yin-bir-zor2004,
ros-ols2015}.  The more significant challenge of extending our work to
three dimensions is guaranteeing that the mesh does not become
distorted, which can be done, for example, by using an extension of the
two-dimensional {\thL} formulation~\cite{amb-sie-tlu2013}.

The second limitation is that the eroding bodies are fixed.  While
mobile eroding bodies have been considered~\cite{Mitchell2016}, only a
single body is simulated.  When multiple bodies are present, especially
in dense suspensions, numerical errors can cause bodies to unphysically
come into contact.  However, by including a small force and
torque~\cite{lu-rah-zor2017}, contact can be avoided.

\paragraph{Related work}
Simulating eroding bodies has been studied analytically, experimentally,
and numerically.  Because of the scales involves with geophysical flows,
most of these works consider high Reynolds number flows.  For instance,
the experimental work of Moore et al.~\cite{moo-ris-chi-zha-she2013}
considers on the order of $10^4$.  For low Reynolds number flows,
Mitchell and Spagnolie~\cite{Mitchell2016} simulate the erosion of a
mobile body in three dimensions, but they only consider single body.

@@

\paragraph{Outline of the paper}
In Section~\ref{s:formulation}, we describe the governing equations for
both the fluid and describe the erosion model using a \thL framework.
In Section~\ref{s:method}, the numerical methods in both space and time
are described.  Section~\ref{s:results} gives several numerical
examples.  Finally, concluding remarks are made in
Section~\ref{s:conclusions}

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[htbp]
\begin{center}
\includegraphics[width = 0.9 \textwidth]{./figs/50bod.pdf}
\caption{\label{fig:50bodies} 50 bodies with vorticity}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Figure generated using 50circ128.in with epsfac = 10, sigfac = 10, dt = 2e-4, fixpdrop = 1.

>>>>

\title{Viscous Erosion in Porous Media}

@@

\begin{keyword}
  Stokes flow \sep Erosion \sep Boundary integral method \sep Shear
  stress \sep \thL~formulation
\end{keyword}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction\label{s:intro}}
Erosion is a fluid-mechanical process that occurs when a fluid alters
the geometry.  Erosion is responsible for a variety of processes, many
of which are geophysical.  For example, sand dunes are longer on the
windward side because of the wind erosion.  Another applications is
erosion caused by groundwater flow, where a slow moving fluid erodes
channels, or conduits, in the earth which creates a heterogeneous media
that has a large effect on the resulting flow.  In both these examples,
there is a feedback process between the flow and the erodible
bodies---the shape of the eroding body affects the flow, while the flow
changes the shape of the eroding body.  This fluid-mechanical feedback
system has been simulated in idealized cases such as a single eroding
two-dimensional body in a uniform flow~\cite{moo-ris-chi-zha-she2013} or
a single mobile eroding three-dimensional body~\cite{mit-spa2016}. To
further develop insight into the effect of two-dimensional erosion, we
consider erosion of a multiply-connected domain
(Figure~\ref{fig:50bodies}), such as a porous media that is typically
used to model groundwater flow.

Because of the scales involved with groundwater flow, we will restrict
our attention to Stokesian dynamics.  For groundwater flow, Darcy's
equation are often used, but this assumes a homogeneous geometry which
is not always a fair assumption.  Instead, we will use the
incompressible Stokes equations to represent the flow, and the
individual pores will be be resolved, resulting in a complex
multiply-connected geometry.  A boundary integral equation is
well-suited as it can resolve complex geometries and the non-negligible
interactions between bodies are resolved, all the while high-order
accuracy and computational efficiency are achieved.  To simulate the
erosion, each body will erode at a rate that is proportional to the
shear stress.  To further visualize the effect of erosion, we compute
the vorticity of the flow (Figure~\ref{fig:50bodies}) which is equal to
the shear stress in the Stokesian limit.  Finally, to evolve the
geometry, a second-order time stepping method that guarantees that the
mesh does not become distorted will be employed.

\paragraph{Contributions} The present study will develop numerical
methods to simulate viscous erosion of multiply-connected
two-dimensional geometries.  By assuming the flow is viscous, we are
assuming the flow is very slow and in small length scales, as is the
case in groundwater flow.  Moreover, by allowing for complex
multiply-connected domains, the effects due to interacting bodies on the
erosion are captured.

@@

\paragraph{Limitations} There are two main limitations of the current
work.  First, the flow is assumed to be two-dimensional.  Although
several flows can be described to good accuracy under a two-dimensional
approximation, certain phenomena occur only in three dimensions.  The
boundary integral equation formulation for the fluid solver exists, and
the main challenges of a BIE solver have been addressed by other groups
including surface representation~\cite{yin-bir-zor2006,
vee-rah-bir-zor2011}, near-singular integration~\cite{yin-bir-zor2006,
kli-tor2016b}, and fast summation methods~\cite{yin-bir-zor2004,
ros-ols2015}.  The more significant challenge of extending our work to
three dimensions is guaranteeing that the mesh does not become
distorted, which can be done, for example, by using an extension of the
two-dimensional {\thL} formulation~\cite{amb-sie-tlu2013}.

The second limitation is that the eroding bodies are fixed.  Mitchell
and Spagnolie~\cite{mit-spa2016} simulate a single mobile eroding body
in three dimensions have been considered~\cite{mit-spa2016}, but
additional challenges are caused when there are multiple bodies. In
particular, especially in dense suspensions, bodies can become extremely
close to one another, and the interaction between the bodies is
difficult to resolve unless the underlying mesh is refined to levels
that are intractable. In future work we plan to consider mobile eroding
bodies, but include a small force and torque so that unphysical contact
is avoided~\cite{lu-rah-zor2017}.

\paragraph{Related work}
Simulating eroding bodies has been studied analytically, experimentally,
and numerically.  Because of the scales involves with geophysical flows,
most of these works consider high Reynolds number flows.  For instance,
the experimental work of Moore et al.~\cite{moo-ris-chi-zha-she2013}
considers on the order of $10^4$.  For low Reynolds number flows,
Mitchell and Spagnolie~\cite{mit-spa2016} simulate the erosion of a
mobile body in three dimensions, but they only consider single body.

@@

\paragraph{Outline of the paper}
In Section~\ref{s:formulation}, we describe the governing equations for
both the fluid and describe the erosion model using a \thL~framework.
In Section~\ref{s:method}, the numerical methods in both space and time
are described.  Section~\ref{s:results} gives several numerical
examples.  Finally, concluding remarks are made in
Section~\ref{s:conclusions}

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[htbp]
\begin{center}
\includegraphics[width = 0.9 \textwidth]{./figs/50bod.pdf}
\caption{\label{fig:50bodies} A time lapse of 50 bodies eroding in a
viscous flow.  The boundary conditions are a constant pressure drop that
creates a flow from the left to right.  The rate of erosion is
proportional to the shear stress which is equivalent to the vorticity
evaluated on the bodies.  The colors represent the vorticity, so regions
with large vorticity (in magnitude) are regions where the flow will
erode the fastest.}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Figure generated using 50circ128.in with epsfac = 10, sigfac = 10, dt = 2e-4, fixpdrop = 1.
